---
title: "LA's BEST R Tutorial"
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
knitr::opts_chunk$set(echo = FALSE)
```

Note to self: run with command rmarkdown::run("R_tutorial.Rmd") (must be in wd)


## Data Frames

**Data frame:** A data frame is a list of vectors which are of equal length. A matrix contains only one type of data, while a data frame accepts different data types (numeric, character, factor, etc.). (guru99)

### Creating a data frame 

Run the following code to create and print a data frame:

```{r dfs, exercise=TRUE, exercise.completion=T}
# Create a, b, c, d variables
a <- c(10,20,30,40)
b <- c('book', 'pen', 'textbook', 'pencil_case')
c <- c(TRUE,FALSE,TRUE,FALSE)
d <- c(2.5, 8, 10, 7)

# Join the variables to create a data frame
df <- data.frame(a,b,c,d)
df
```

### Naming columns 

The following code renames the columns as:
1. a <- cat
2. b <- dog
3. c <- mouse
4. d <- sheep

Edit the code so that the names make sense with the data:

```{r df_create, include=FALSE}
# Create a, b, c, d variables
a <- c(10,20,30,40)
b <- c('book', 'pen', 'textbook', 'pencil_case')
c <- c(TRUE,FALSE,TRUE,FALSE)
d <- c(2.5, 8, 10, 7)

# Join the variables to create a data frame
df <- data.frame(a,b,c,d)
```

```{r dfs_name, exercise=TRUE, exercise.setup="df_create"}
names(df) <- c("cat", "dog", "mouse", "sheep")
df
```

```{r dfs_name-solution}
names(df) <- c("ID", "items", "in-store", "price")
df
```


### Selecting rows and columns

We can choose to only use certain portions of a data frame by selecting rows and columns. Using matrix notation, you can subset the data framee `df` with the code `df[1,2]` which will select row 1 and column 2 of the data frame.

Print out the third row of the price column:
```{r df_create2, include=FALSE}
# Create a, b, c, d variables
a <- c(10,20,30,40)
b <- c('book', 'pen', 'textbook', 'pencil_case')
c <- c(TRUE,FALSE,TRUE,FALSE)
d <- c(2.5, 8, 10, 7)

# Join the variables to create a data frame
df <- data.frame(a,b,c,d)

names(df) <- c("ID", "items", "in-store", "price")
```

```{r object_select, exercise=TRUE, exercise.setup="df_create2"}

```

```{r object_select-solution}
df[3, 3]
```

You can also select the first entire row with `df[1, ]` or the entire first column with `df[ , 1]`. You can also select columns by their name and the `$` notation.

Change the following code to select the items column:
```{r df_create3, include=FALSE}
# Create a, b, c, d variables
a <- c(10,20,30,40)
b <- c('book', 'pen', 'textbook', 'pencil_case')
c <- c(TRUE,FALSE,TRUE,FALSE)
d <- c(2.5, 8, 10, 7)

# Join the variables to create a data frame
df <- data.frame(a,b,c,d)

names(df) <- c("ID", "items", "in-store", "price")
```

```{r column_select, exercise=TRUE, exercise.setup="df_create3"}
df$price
```

```{r column_select-solution}
df$items
```


### Append a column

After we create a data frame, we can add a new column. We do this by creating a vector and then appending it to the end of the data frame.

The following code creates a vector called "quantity" and adds it to the data frame as a column. Add to this code to create a vector called "student" which contains true/false values:

```{r df_create4, include=FALSE}
# Create a, b, c, d variables
a <- c(10,20,30,40)
b <- c('book', 'pen', 'textbook', 'pencil_case')
c <- c(TRUE,FALSE,TRUE,FALSE)
d <- c(2.5, 8, 10, 7)

# Join the variables to create a data frame
df <- data.frame(a,b,c,d)

names(df) <- c("ID", "items", "in-store", "price")
```

```{r newcolumn, exercise=TRUE, exercise.eval=TRUE, exercise.setup="df_create4"}
# Create a new vector
quantity <- c(10, 35, 40, 5)

# Add `quantity` to the `df` data frame
df$quantity <- quantity
df
```

```{r newcolumn-solution}
# Create a new vector
quantity <- c(10, 35, 40, 5)

# Add `quantity` to the `df` data frame
df$quantity <- quantity
df

student <- c(T, F, T, F)
df$student <- student
df
```

### Quiz

```{r quiz}
quiz(
  question("Choose all that apply. How can I select the second column (named bp) of a data frame?",
    answer("df$bp", correct=T),
    answer("df[2, ]"),
    answer("df[ , 2]", correct=T),
    answer("df[2, 2]")
  )
)
```

